---
title: "R-Ladies Gainesvile"
subtitle: "Intro. to blogdown"
author: "<br><br>Justin Millar <br><br>`r anicon::faa('twitter', animate='float',rtext='@justinjmillar')`  "
date: "<br>2018-10-23<br>"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["kunoichi", "ninjutsu" , "assets/ninpo.css"]
    self_contained: false
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

class: middle center inverse bg-main1

# Everyone leaves with a working website

--

## (hopefully)

---

layout: false
class: split-33 with-thick-border

.column.bg-main1[.content.vmiddle[

# What is blogdown? 
]]
.column[.content[

## - Blogdown is a package in R used to make "static" websites
<br><br>
## - It does not require knowledge of HTML/CSS/JavaScript etc.
<br><br>
## - Use a program call Hugo to create HTML files from Markdown (.md) and RMarkdown (.Rmd) files
<br><br>
## We'll rely on [this great book to get started with blogdown](https://bookdown.org/yihui/blogdown/).
]]

---

class: middle inverse bg-main1

# Getting started 

### First we'll install the package:
```{r getting-started, eval=F}
# Install from CRAN
install.packages("blogdown")

# Install development verison from GitHub
devtools::install_github("rstudio/blogdown")
```

### Then we install Hugo using the `blogdown` package
```{r install-hugo, eval=FALSE}
blogdown::install_hugo()
```

---

class: middle inverse bg-teal

# A mininal website

```{r new-site, eval=FALSE}
blogdown::new_site()
```


#### Blogdown should have made a bunch of files/folders, and your website should be displayed in the Viewer tab.

### Three key files/directories:

### - `config.toml`

### - `content`

### - `public`


---

layout: false
class: split-33 with-thick-border

.column.bg-main1[.content.vmiddle[

# The configuration file: `config.toml` 
]]

.column[.content[
### The `config` file is used to to set the global configurations of your site:

```
baseurl = "/"
languageCode = "en-us"
title = "A Hugo website"
theme = "hugo-lithium"

[[menu.main]]
    name = "About"
    url = "/about/"
[[menu.main]]
    name = "GitHub"
    url = "https://github.com/rstudio/blogdown"
[[menu.main]]
    name = "Twitter"
    url = "https://twitter.com/rstudio"
```
### Check out [this chapter](https://bookdown.org/yihui/blogdown/configuration.html) for more info on this file.

]]

---

layout: false
class: split-33 with-thick-border

.column.bg-blue[.content.vmiddle[

# The content folder 
]]

.column[.content[
### The `content` folder contains all of the Markdown and RMarkdown files that will make up the information in our pages:

```
├── _index.md
├── about.md
├── vitae.md
├── post/
│   ├── 2017-01-01-foo.Rmd
│   ├── 2017-01-02-bar.md
│   └── ...
└── ...
```

### These cheatsheets for using Markdown and [RMarkdown](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf) are really helpful.

]]

---

layout: false
class: split-33 with-thick-border

.column.bg-red[.content.vmiddle[

# The public folder 
]]

.column[.content[
### The `public` folder is where Hugo does it's magic, taking using the `config` file and `content` folder to create your website.

<br><br>

### If you just want to look at your site locally (as in, the files located on your computer), you'll want to open the `index.html` file in the public folder using a web browser.

<br><br>

### Later we will see different ways of using the public folder to host your website online.

]]

---

layout: false
class: split-33 with-thick-border

.column.bg-orange[.content.vmiddle[

# Custom themes 
]]

.column[.content[

### There are many different themes Hugo can use to create your site:
<br><br>
## [https://themes.gohugo.io/](https://themes.gohugo.io/)
### In my experience, the easist way to use a new theme is to start from scratch. Find a theme from the link above and pass the GitHub repository name to blogdown:

```{r new-theme, eval=FALSE}
blogdown::new_site(theme = "gcushen/hugo-academic")
```


### You can also use `blogdown::install_theme` to add a theme to an existing site, but this may screw up you config file.

<br><br>

### Also, it's best to start off with a simpler theme in the beginning, which will be easier to convert down the line.

]]

---

layout: false
class: split-25 with-thick-border

.column.bg-indigo[.content.vmiddle[

# Writing a new post/page 
]]

.column[.content[

### You'll notice that `blogdown` added some useful RStudio Addins, including "New Post" which runs:
```{r new-post, eval=FALSE}
blogdown::new_post_addin()
```

### Running this function will launch a Shiny app were we can specific some information about out posts, which will get placed in the YAML:

```
---
title: Example post
author: My Name
date: '2018-10-23'
slug: example-post
categories:
  - R
tags:
  - R Markdown
---
```
### This is also where we can choose between an RMarkdown file (.Rmd) and a regular Markdown file (.md)

]]

---


layout: false
class: split-25 with-thick-border

.column.bg-green[.content.vmiddle[

# Deploying your site
]]

.column[.content[

### Up to this point, our website only exists on our computer. There are [many options for deploying your site](https://bookdown.org/yihui/blogdown/deployment.html), but one of the easiest (and free) ways is to use Netlify

<br><br>

### We'll create a free account (can use Google, Github, etc.), then drag the `public` folder into the Deploy box.
<br><br>

### Next we can change the Domain to something more appropriate under Domain Settings
<br><br>
### Note that we'll have to re-drag the `public` folder each time we a change to the site. If you're familiar with git and GitHub, it's better to keep the site in a GitHub repo and point Netlify to the `public` folder. Netlify will automatically reconstruct the site when ever you make a push to the repo. 

]]

---
layout: false
class: split-25 with-thick-border

.column.bg-amber[.content.vmiddle[

# Using the static folder

]]

.column[.content[

### You can also place files in the static folder, which will then be accessible using your URL. For example, I can place `example-data.csv` in the static folder and then that file can be downloaded at www.my-site.com/example-data.csv. 

<br><br>

### WARNING: Hugo will place the files in `static` and place it into `public` when it creates the site, however it has some funny tendencies to overwrite files.

### This means you should be **extra careful** that there are not similarly named files in the `content` and `static` folders.
<br><br>
### I've also have some issues with certain themes with the `static` folder, so it may take some troubleshooting

]]

---

class: middle center bg-main1

# Thanks!!


